---
title: "Greenhouse Gas Reporting Program Data"
author: "Laura Cech and Andrea Duran Liceaga"
date: "2023-Fall"
output: html_document
---

Intro: Given the ambitious goals of Maryland officials to reduce greenhouse gas emissions (by 60 percent by 2031), we propose to examine EPA data to answer fundamental questions:

1. How do greenhouse gas emissions in MD compare/ rank nationally? 
2. What are the locations of top greenhouse gas emitters?
3. How do the hotspots of emissions match with sites being supervised by state and federal environmental authorities? 
4. What are the demographics (median income, race) in areas with the most greenhouse emissions?
5. What industries/ companies emit the most and what are their parent companies? 

We believe these questions are newsworthy because they address some of the things people might want to know about how our state is doing regarding emissions.

Datasets come from: https://www.epa.gov/ghgreporting/data-sets

 
Deliverable 2:
Loads and cleans the core data set to prepare for analysis. 

We have loaded national and state data from EPA and census data. Based on our questions, we believe we have the data needed. We also may be interested in exploring additonal features of the FLIGHT dashboard.

Describe the limits of the data in terms of timeframe, what is and is not included and any codes or other information you'd need to find definitions for, and what you've done to better define it.
-- https://www.epa.gov/ghgreporting/resources-subpart-ghg-reporting shows the codes of the subparts
-- We found a good glossary (https://www.epa.gov/energy/greenhouse-gas-equivalencies-calculator) that translates the amount of emissions into real world effects.

Deliverable 3:
At a minimum, your data notebook should load the data you'll use with this project and have demonstrated work on at least two of the questions you are seeking to answer. 

Deliverable 4:
Shows a draft of data notebook in which you demonstrate substantial progress on the questions you set out to answer at the beginning of the project. This means you should have started on each of them, but you may not have finished.
Has narrative for your code that explains each of your questions, your findings so far and what you do and don't know as a result of them.

Identifies your most newsworthy finding so far and explain why it is.
We think we it's newsworthy and worth investigating the clump of superfund sites in Cecil County, centered around Elkton, Maryland. We also wonder why there was a dip in emissions in 2017. (The dip during covid made sense.)

## Load libraries and establish settings

#IMPORT

```{r}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(janitor)
library(tigris)
library(tidycensus)
library(dplyr)
library(ggplot2)
library(mapview)
library(ggthemes)
#asked chatgpt for code to load one sheet of a spreadsheet
library(readxl)

```


```{r}

#This is the data from FLIGHT to look at emissions from the state of Maryland
md_emissions2022 <- read_excel("data/md_emissions_2022.xls", skip = 5)
glimpse(md_emissions2022)

md_emissions2021 <- read_excel("data/flight.xls", sheet = "2021", skip = 6) 

md_emissions2020 <- read_excel("data/flight.xls", sheet = "2020", skip = 6) 

md_emissions2019 <- read_excel("data/flight.xls", sheet = "2019", skip = 6) 

md_emissions2018 <- read_excel("data/flight.xls", sheet = "2018", skip = 6) 

md_emissions2017 <- read_excel("data/flight.xls", sheet = "2017", skip = 6) 

md_emissions2016 <- read_excel("data/flight.xls", sheet = "2016", skip = 6) 

md_emissions2015 <- read_excel("data/flight.xls", sheet = "2015", skip = 6) 

md_emissions2014 <- read_excel("data/flight.xls", sheet = "2014", skip = 6) 

md_emissions2013 <- read_excel("data/flight.xls", sheet = "2013", skip = 6) 

md_emissions2012 <- read_excel("data/flight.xls", sheet = "2012", skip = 6) 

md_emissions2011 <- read_excel("data/flight.xls", sheet = "2011", skip = 6) 

md_emissions2010 <- read_excel("data/flight.xls", sheet = "2010", skip = 6) 

#this is getting the data from the Census API file for median income from 2021 by County in the state of Maryland.
census_api_key("77872777955b38b9334cbbae8eea25f316159387")

md_county_income <- get_acs(geography = "county",
              variables = c(medincome = "B19013_001"),
              year = 2021,
              state = "MD") |> 
  mutate(min = estimate - moe, max = estimate + moe) |> 
  glimpse()

superfund_fixed <- read_csv("data/superfund-fixed.csv", skip = 10)

```

# CLEAN and load all years of data. We made sure all values are consistent including addressing the Baltimore County/ Baltimore City issue and capitalization. We've removed duplicates.
```{r}

md_emissions2022 <- md_emissions2022 |> 
  #to view largest to smallest quantity
  clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2010 <- md_emissions2010 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2011 <- md_emissions2011 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2012 <- md_emissions2012 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2013 <- md_emissions2013 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2014 <- md_emissions2014 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2015 <- md_emissions2015 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2016 <- md_emissions2016 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2017 <- md_emissions2017 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2018 <- md_emissions2018 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2019 <- md_emissions2019 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2020 <- md_emissions2020 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

md_emissions2021 <- md_emissions2021 |> 
   clean_names() |> 
  rename(ghg_quantity = ghg_quantity_metric_tons_co2e) |>
  mutate(county_name = str_to_title(str_remove_all(county_name,"COUNTY"))) |>
  mutate(county_name = str_trim(county_name,side="both")) |>
  #mutate(county_name = str_to_title(str_remove_all(county_name," CITY"))) |> 
  mutate(county_name = str_to_upper(county_name)) |> 
  mutate(facility_name = str_to_upper(facility_name)) |> 
  mutate(reported_address = str_to_upper(reported_address)) |>
  mutate(city_name = str_to_upper(city_name)) |>
  mutate(parent_companies = str_to_upper(parent_companies)) |>
  distinct()

emissions_grouped_by_county <- md_emissions2022 |>
  group_by(county_name) |> 
  summarise(total_emissions = sum(ghg_quantity)) 

md_county_income <- md_county_income |> clean_names() |> 
  mutate(name = str_to_title(str_remove_all(name,"County, Maryland"))) |>
  mutate(name = str_to_title(str_remove_all(name, ", Maryland"))) |> 
  mutate(name = str_trim(name,side="both")) |>
  mutate(name = str_to_upper(name))

md_2022_emissions_with_income <- md_county_income |>
  rename(county_name = name) |> 
  left_join(emissions_grouped_by_county)

superfund_fixed_filtered <- superfund_fixed |> 
  clean_names() 

```

```{r}
#to compare Md and the rest of the nation
#read in summary file from EPA
national_emissions2022 <- read_excel("2022_data_summary_spreadsheets/ghgp_data_by_year_2022.xlsx", sheet = "Direct Emitters", skip = 3) 
glimpse(national_emissions2022)
```

#ANALYSIS


#1. How do greenhouse gas emissions in MD compare/ rank nationally? 

Maryland is 37th in the ranking of all states and territories with Texas having the most direct emissions in 2022 and total facilities emitting greenhouse gas and DC having the least emissions and the Virgin Islands having the fewest facilities.


```{r}

#group by state and count
national_filtered <- national_emissions2022 |>
   group_by(State) |>  
  summarise(count=n()) |>
  arrange(desc(count))
```

```{r}
#chat gpt helped troubleshoot error
#we chose direct emissions because it provided the most accurate and full picture
total_emissions_per_state <- national_emissions2022 %>%
  group_by(State) %>%
  summarise(total_emissions = sum(`2022 Total reported direct emissions`, na.rm = TRUE)) |>
  arrange(desc(total_emissions))
```

#2. What are the locations of top greenhouse gas emitters in Maryland and how has that changed in the last 12 years?

https://datawrapper.dwcdn.net/BYVo2/1/

Overall, Greenhouse Gas emissions have declined since 2010. The least amount of emissions were produced in 2020 (which coincides with the pandemic), and in 2017. It's unclear why in 2017 the total amount of emissions were so low. 

There were 67 total facilities emitting greenhouse gas in 2012. A decade later in 2022, there were 52 total facilities emitting greenhouse gas, according to the EPA.

The top emitters in 2012 were Brandon Shores LLC, Morgantown and Chalk Point.
Whereas in 2022, the top emitters were "BRANDON SHORES LLC", "LEHIGH CEMENT COMPANY LLC" and "KEYS ENERGY CENTER." A decade later, and Brandon Shores LLC (a power plant) is still the top emitter. In 2022, it emitted 2,219,520 metric tons of Greenhouse Gases and in 2012, it emitted 4,854,518 metric tons (rounded to nearest whole number), so that gives us a 54% drop in emissions.

A couple of the most interesting finds were University of Maryland that came in 18th out of 52 and Holcim (US), Inc. which is a building material manufacturer that is supposed to be a "global leader in sustainable construction" and works to "make greener infrastructure" but they are the 9th GHG emitter. The counties that have no emissions are Queen Anne's, Somerset, St. Mary's and Kent counties.

In Charles county, where direct emissions totaled 2,843,429 metric tons for 2022, it would take 791 wind turbines running for a year to offset. Source: https://www.epa.gov/energy/greenhouse-gas-equivalencies-calculator

```{r}
locations_md <- md_emissions2022 |> 
  arrange(desc(ghg_quantity)) |> 
  mutate(zip_code = as.character(zip_code)) |> 
  group_by(county_name) |> 
  summarise(sum(ghg_quantity))

counties <- counties()
```

```{r}
#calculating most emisions and arranging to compare 2022 and a decade before 2012
md_emissions2022 |> 
  arrange(desc(ghg_quantity))


md_emissions2012 |> 
  arrange(desc(total_reported_direct_emissions))

```

```{r}
#analyzing all years

#total number of emissions to join by row
totals_2022 <- md_emissions2022 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2022")
totals_2021 <- md_emissions2021 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE)) |>  
             mutate(year = "2021")
totals_2020 <- md_emissions2020 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2020")
totals_2019 <- md_emissions2019 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2019")
totals_2018 <- md_emissions2018 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2018")
totals_2017 <- md_emissions2017 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2017")
totals_2016 <- md_emissions2016 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2016")
totals_2015 <- md_emissions2015 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2015")
totals_2014 <- md_emissions2014 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2014")
totals_2013 <- md_emissions2013 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2013")
totals_2012 <- md_emissions2012 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2012")
totals_2011 <- md_emissions2011 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2011")
totals_2010 <- md_emissions2010 |> 
  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(year = "2010")

combined_years <- bind_rows(totals_2022, totals_2021, totals_2020, totals_2019, totals_2018, totals_2017, totals_2016, totals_2015, totals_2014, totals_2013, totals_2012, totals_2011, totals_2010)

write_csv(combined_years,"data/combined_years.csv")

```

```{r}
#writing a file to use for datawrapper
combined_emissions <- bind_rows(md_emissions2022, md_emissions2021)

write_csv(combined_emissions,"data/combined_emissions1.csv")
```


#3. How do the hotspots of emissions match with sites being supervised by state and federal environmental authorities? 

The most superfunds in Maryland are located in Cecil County. Interestingly enough, three of those superfunds in Cecil County are in the town of Elkton. We suspect there's a good story here about why this is true. In addition to groundwater contamination from an old fireworks production facility, we see solvent recycling operations and industrial waste disposal activities took place at a former sand and gravel quarry.

```{r}
#analyzing superfund to show those in the final list of sites being supervised by environmental officials.
superfund_fixed_filtered |> 
  filter(str_detect(npl_status, "Final NPL")) |> 
  group_by(county) |> 
  arrange(desc(county))

superfund_fixed_filtered |> 
  filter(str_detect(npl_status, "Final NPL")) |> 
  group_by(county) |> 
  summarise(count = n()) |> 
  arrange(desc(count))

```


#4. What are the demographics (median income) in areas with the most greenhouse emissions in Maryland?

The median income of the counties with the highest emissions have among the highest median incomes in Maryland.
In Charles County, which has the most emissions, median income is about $109,905. Anne Arundel County's median income is about $109,958 and Prince George's is about $92,513.

 
```{r}
#we arranged from most to least to see how the variable median income would affect the data
md_2022_emissions_with_income |> 
  arrange(desc(total_emissions))


write_csv(md_2022_emissions_with_income, "data/md_2022_emissions_with_income.csv")  

```

```{r}
#this wasn't the best visualization but we made others
ggplot(md_2022_emissions_with_income, aes(x = max, y = total_emissions)) +
  geom_point(color = "blue", size = 3) +
  labs(title = "Total emissions and median income", x = "Income", y = "GHG") +
  geom_text(aes(label = county_name), check_overlap = TRUE)

```


```{r}
#we made a map visualization of median income and total emissions per county using mapview
md_counties <- counties |>
  filter(STATEFP == "24") |>
  clean_names() |>
  arrange(namelsad)


locations_map <- md_counties |>
  arrange() |>
  bind_cols(md_2022_emissions_with_income)

mapview(locations_map, zcol="total_emissions")

```



#5. What industries emit the most greenhouse gases in Maryland? 
General Stationary Fuel Combustion Sources= 16,384,801
Municipal Solid Waste Landfills= 884,420
Petroleum and Natural Gas Systems= 1,466,486
Electricity Generation= 10,636,599
Suppliers of Natural Gas/Liquids= 177,591
Suppliers of Carbon Dioxide= 1,406,085
Cement Production= 3,366,799

See our chart at: https://datawrapper.dwcdn.net/unfwg/1/

In Maryland, the fuel (42%), electricity (27%) and cement production (17%) industries produced the most emissions in 2022.


```{r}
#we filtered all data for each industry, joined all results by totals and wrote a csv file to use datawrapper
# https://www.epa.gov/ghgreporting/resources-subpart-ghg-reporting shows the codes of the subparts
Fuel<- md_emissions2022 |>
  filter(str_detect(subparts, "C")) 
 

```


```{r}
Landfills<- md_emissions2022 |>
  filter(str_detect(subparts, "HH")) 
  

```

```{r}
Petsystems<- md_emissions2022 |>
  filter(str_detect(subparts, "W"))
 
```

```{r}
Electricity<- md_emissions2022 |>
  filter(str_detect(subparts, "D")) 
 
```

```{r}
carbon<- md_emissions2022 |>
  filter(str_detect(subparts, "PP")) 
```


```{r}
cement<- md_emissions2022 |>
  filter(str_detect(subparts, "H")) 
 
```


```{r}
#join all results by totals
total_fuel <- Fuel  |>  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(source = "fuel")
total_landfills <- Landfills  |>  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(source = "landfill")
total_petsystems <- Petsystems  |> summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(source = "pet systems")
total_electricity <- Electricity  |>  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(source = "electricity")
total_carbon <- carbon |>  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(source = "carbon")
total_cement <- cement  |>  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(source = "cement")
total_natgas <-Natgas  |>  summarise(total_ghg = sum(ghg_quantity, na.rm = TRUE))  |>  
             mutate(source = "natural gas")

combined_totals <- bind_rows(total_fuel, total_landfills, total_petsystems, total_electricity, total_carbon, total_cement, total_natgas)

```


```{r}
write_csv(combined_totals, "data/combined_totals.csv")  
```

